### 2011-01-12 paper/abstract.tex
Added copy of report as paper

This paper presents an array of approaches to optimizing a quadrupedal
gait for forward speed.  We implement, test, and compare different
learning strategies including uniform and Gaussian random hill
climbing, policy gradient reinforcement learning\cite{kohl},
Nelder-Mead simplex\cite{nm}, new predictive methods based on linear
and support vector regression, and an evolved neural network
(HyperNEAT)\cite{2}.  We compare results to a baseline random search
method.  Many of the methods resulted in walks significantly faster
than previously hand-tuned gaits.

@@

\edit{fix citations}

>>>>

This paper presents an array of approaches to optimizing a quadrupedal
gait for forward speed.  We implement, test, and compare different
learning strategies including uniform and Gaussian random hill
climbing, policy gradient reinforcement learning\cite{kohl},
Nelder-Mead simplex\cite{nm}, new predictive methods based on linear
and support vector regression, and an evolved neural network
(HyperNEAT)\cite{clune}.  We compare results to a baseline random
search method.  Many of the methods resulted in walks significantly
faster than previously hand-tuned gaits.

@@

### 2011-01-12 paper/report.tex
Added copy of report as paper

The output each of the learning algorithms is a function of time,
$f(t)$, that outputs a vector of commanded motor positions.  This
function is generated using a parametrized motion model, described in
\secref{motion}.

@@

\item \emph{Nelder-Mead simplex method}\cite{nm}: The Nelder-Mead
  simplex method creates an initial simplex with 6 vertices. The
  initial parameter vector is stored as the first vertex and the other
  five vertices are created by adding to one dimension at a time one
  tenth of the allowable range for that parameter. It then tests the
  fitness of each vertex and based on these fitnesses, it reflects the
  worst point over the centroid in an attempt to improve it.  However,
  to prevent cycles and becoming stuck in local minima, several other
  rules are used.  In general, the worst vertex is reflected over the
  centroid. If the reflected point is better than the second worst
  point and worse than the best point, then the reflected point
  replaces the worst. If the reflected point is better than the best
  point, the simplex is expanded in the direction of the reflected
  point. The better of the reflected and the expanded point replaces
  the worst point. If the reflected point is worse than the second
  worst point, then the simplex is contracted away from the reflected
  point. If the contracted point is better than the reflected point,
  the contracted point replaces the worst point. If the contracted
  point is worse than the reflected point, the entire simplex is
  shrunk \cite{NMWebsite}.

@@

\item \emph{SVM regression}: Similarly to linear regression, this
  model starts with several random vectors, but this time they are
  chosen in a small neighborhood about some initial random vector.
  These vectors (generally 8) are evaluated, and a support vector
  regression model is fit to the observed fitnesses.  To choose the
  next vector for evaluation, we randomly generate some number
  (typically 100) of vectors in the neighborhood of the best observed
  gait, and select for evaluation the vector with the best predicted
  performance.  We suspected that if we always chose the best
  predicted point out of 100, we may end up progressing along a narrow
  subspace, prohibiting learning of the true local fitness function.
  Alternately, we would always choosing exploitation of knowledge
  vs. exploration of the space.  To address this concern, we added a
  parameter dubbed \code{bumpBy} that added noise to the final
  selected point before it was evaluated.

  Such a method naturally has many tunable parameters, and we
  endeavored to select these parameteres by tuning the method in
  simulation.  To estimate the performance of the algorithm, we ran it
  agains a simulation with a known optimum.  The simulated function
  was in the same five dimensional parameter space, and simply
  returned a fitness determined as the height of a Gaussian with a
  random mean.  The width of the Gaussian in each dimension was 20\%
  of the range of each dimension, and the maximum value at the peak
  was 100.  \figref{svm_sim_results} shows the learning results on
  this simulated model using the ultimately selected SVM parameters.
  Interestingly, a non-zero value of \code{bumpBy} resulted in better
  learning than a zero valued, noise free (exploration free) learning.

  Ultimately, however, the version of SVM tuned for simulation did not
  yet show competitive performance on the real robot.  We tried tuning
  some parameters on the real robot, but after some amount of tuning,
  the method still exibited too little exploration and easily became
  stuck in local minima.

\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{clune}}: Near
  the end of term, we hacked together an interface between HyperNEAT
  and the robot, requiring a slightly modified strategy interface.
  Preliminary HyperNEAT runs were promising and resulted in serveral
  interesting gaits.  Unfortunately, the gaits generated by HyperNEAT
  also tended to stress the robot more than typical gaits had before,
  and the servos would often overheat and malfunction, requiring
  restart.  We think these issues may be addressed by adding a small
  layer between the HyperNEAT strategy and the robot that disallows
  quickly shifting commanded positions, and we hope to be able to test
  these methods further once this filter is in place.

@@

In addition to running strategies to optimize parameterized gaits, we
also wanted to investigate the space of possible gaits.  To accomplish
this, we selected a parameter vector that resulted in motion, but not
an exceptional gait, and plotted performance along each dimension
individually.  In addition, we duplicated each measurement to be able
to estimate the measurement noise at each point.  Results for this
exploration are shown in \figref{explore_dim_1} through
\figref{explore_dim_5}.

\figvarp{explore_dim_1}{.5}{Motion for varying just dimension 1}{}

\figvarp{explore_dim_2}{.5}{Motion for varying just dimension 2}{}

\figvarp{explore_dim_3}{.5}{Motion for varying just dimension 3}{}

\figvarp{explore_dim_4}{.5}{Motion for varying just dimension 4}{}

\figvarp{explore_dim_5}{.5}{Motion for varying just dimension 5}{}

>>>>

The output each of the learning algorithms is a function of time,
$f(t)$, that outputs a vector of commanded motor positions.  This
function is generated using a parametrized motion model, described in
\secref{implement}.

@@

\item \emph{Nelder-Mead simplex method}\cite{nm}: The Nelder-Mead
  simplex method creates an initial simplex with 6 vertices. The
  initial parameter vector is stored as the first vertex and the other
  five vertices are created by adding to one dimension at a time one
  tenth of the allowable range for that parameter. It then tests the
  fitness of each vertex and based on these fitnesses, it reflects the
  worst point over the centroid in an attempt to improve it.  However,
  to prevent cycles and becoming stuck in local minima, several other
  rules are used.  In general, the worst vertex is reflected over the
  centroid. If the reflected point is better than the second worst
  point and worse than the best point, then the reflected point
  replaces the worst. If the reflected point is better than the best
  point, the simplex is expanded in the direction of the reflected
  point. The better of the reflected and the expanded point replaces
  the worst point. If the reflected point is worse than the second
  worst point, then the simplex is contracted away from the reflected
  point. If the contracted point is better than the reflected point,
  the contracted point replaces the worst point. If the contracted
  point is worse than the reflected point, the entire simplex is
  shrunk \cite{nm}.

@@

\item \emph{SVM regression}: Similarly to linear regression, this
  model starts with several random vectors, but this time they are
  chosen in a small neighborhood about some initial random vector.
  These vectors (generally 8) are evaluated, and a support vector
  regression model is fit to the observed fitnesses.  To choose the
  next vector for evaluation, we randomly generate some number
  (typically 100) of vectors in the neighborhood of the best observed
  gait, and select for evaluation the vector with the best predicted
  performance.  We suspected that if we always chose the best
  predicted point out of 100, we may end up progressing along a narrow
  subspace, prohibiting learning of the true local fitness function.
  Put another way, we would always choose exploitation of knowledge
  vs. exploration of the space.  To address this concern, we added a
  parameter dubbed \code{bumpBy} that added noise to the final
  selected point before it was evaluated.

  Such a method naturally has many tunable parameters, and we
  endeavored to select these parameters by tuning the method in
  simulation.  To estimate the performance of the algorithm, we ran it
  against a simulation with a known optimum.  The simulated function
  was in the same five dimensional parameter space, and simply
  returned a fitness determined as the height of a Gaussian with a
  random mean.  The width of the Gaussian in each dimension was 20\%
  of the range of each dimension, and the maximum value at the peak
  was 100.  \figref{svm_sim_results} shows the learning results on
  this simulated model using the ultimately selected SVM parameters.
  Interestingly, a non-zero value of \code{bumpBy} resulted in better
  learning than noise free (exploration free) learning.

  Ultimately, however, the version of SVM tuned for simulation still
  did not show competitive performance on the real robot.  We tried
  tuning some parameters on the real robot, but after some amount of
  tuning, the method still exhibited too little exploration and easily
  became stuck in local minima.

\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{clune}}: Near
  the end of term, we hacked together an interface between HyperNEAT
  -- an implementation of a method for evolving neural networks -- and
  the robot, requiring a slightly modified strategy interface.
  Preliminary HyperNEAT runs were promising and resulted in several
  interesting gaits.  Unfortunately, the gaits generated by HyperNEAT
  also tended to stress the robot more than typical gaits had before,
  and the servos would often overheat and malfunction, requiring
  restarts.  We think these issues may be addressed by adding a small
  layer between the HyperNEAT strategy and the robot that disallows
  quickly shifting commanded positions, and we hope to be able to test
  these methods further once this filter is in place.

@@

In addition to running strategies to optimize parameterized gaits, we
also wanted to investigate the space of possible gaits.  To accomplish
this, we selected a parameter vector that resulted in motion, but not
an exceptional gait, and plotted performance along each dimension
individually.  In addition, we duplicated each measurement to be able
to estimate the measurement noise at each point.  Results for this
exploration are shown in \figref{explore_dim_1} through
\figref{explore_dim_5}.  As is shown in the figures, some dimensions
are smoother than others, and some measurements are fairly noisy.

\figvarp{explore_dim_1}{.65}{Fitness mean and standard deviation
  vs. dimension 1.  The circle is a common point in
  \figref{explore_dim_1} through \figref{explore_dim_5}}{}
\figvarp{explore_dim_2}{.65}{Fitness mean and standard deviation
  vs. dimension 2.  The circle is a common point in
  \figref{explore_dim_1} through \figref{explore_dim_5}}{}
\figvarp{explore_dim_3}{.65}{Fitness mean and standard deviation
  vs. dimension 3.  The circle is a common point in
  \figref{explore_dim_1} through \figref{explore_dim_5}}{}
\figvarp{explore_dim_4}{.65}{Fitness mean and standard deviation
  vs. dimension 4.  The circle is a common point in
  \figref{explore_dim_1} through \figref{explore_dim_5}}{}
\figvarp{explore_dim_5}{.65}{Fitness mean and standard deviation
  vs. dimension 5.  The circle is a common point in
  \figref{explore_dim_1} through \figref{explore_dim_5}}{}
